Repairing cooperation through a cognitive intervention in the repeated Trust Game

### 2023-08-14 article.Rmd
cleaning up MS edits

The strategy of the computerised investor was modelled on behaviour of human investors in a 10-round RTG.<!-- reference to data -->.  Using this data, we estimated a hidden Markov model (HMM) on investors' behaviour with three latent states, reflecting "low-trust", "medium-trust", or "high-trust". Each latent state was associated with a distribution over all possible investor actions (from 0 to 20 investment) that reflected the amount of trust (Figure \ref{fig:HMMPanels}.A). Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of the net return in the previous round (see Figure \ref{fig:HMMPanels}.B).

In order to probe efforts to repair trust, the computerised agent was programmed to provide a low return on round 12 pre-intervention or 13 post-intervention. On all other rounds, the agent's actions were derived from the HMM (disregarding the participant's response on round 12).

@@

Participants played 7 rounds of a repeated Prisoner's Dilemma. In each round, participants could choose a cooperative action with a a reward of 5 if the player also cooperated and the a reward of 1 if not, or a non-cooperative action that would yield a 7 points if the other person chooses the cooperative action and 2 points if not. The Nash equilibrium for a single-round version is to choose the non-cooperative action.

The computerised agent was programmed to act according to a tit-For-tat strategy [@axelrod_evolution_1981], starting with a cooperative action and then mirroring what the other player chose in the preceding round. On round 4, the agent was pre-programmed to choose the defect action, regardless of the participant's preceding action.

@@

The intervention was built on interventions from DBT skills training, asking patients to reflect on the consequences of actions taking in emotional states [@linehan_dbt_2015]. Specifically, participants were presented with a low investment and asked to indicate their response. They were then invited to consider what their ultimate aim in the game was and whether this response was most likely to achieve their aim. The intervention is detailed in the supplementary information.

@@

At the start of the experiment (Figure \ref{fig:timeline}), participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Phase one was a 15 round repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor, participants were asked to provide feedback, with participants in the Intervention condition rating their feeling in terms of valence (from negative to positive) and arousal (from low to high), and participants in the Control condition rating the investment in terms of speed (from slow to fast) and magnitude (from low to high). These ratings were made by indicating a point on a two-dimensional grid. Participants then decided on how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, selfish, trustworthy and friendly they perceived their interaction partner (all on a scale from 1 to 10). After phase one, participants in the intervention condition completed the intervention, and participants in the control condition solved anagrams. Subsequent phase two was similar to phase one, with participants being told they would face a new player. Phase three consisted of 7 rounds of the repeated prisoner's dilemma game (RPD), with participants informed they would face a third player. Participants then completed questionnaires related to mentalising abilities (RFQ8), emotion regulation (DERS), and BPD traits (PAI-BOR).<!-- MS: need to include references for these scales --> They were then asked about the strategy in the games, as well as whether they thought the other players were human or computer agents. They were then debriefed and thanked for their participation.

@@

<!-- Our investment levels are clearly comparable. Our random matching condition is the most directly comparable to Berg et al., (1995) and Charness et al., (2008), both of whom used random matching. Our investment averages 4.97 out of 10 in the individual feedback condition, which falls right in between the investment level of 5.16 documented by Berg et al., (1995) and the 3.73 documented by Charness et al., (2008). The amount returned in the individual feedback condition with random matching is 5.61, which is slightly higher than the amounts returned in Berg et al., (1995) of 3.73 and Charness et al., (2008) of 4.65. The 19% return rate is comparable to results in similar designs. In an almost identical design, Fiedler et al., (2011) report a 17% return rate by responders. This is similar to our finding.  -->

@@

<!--
Estimated marginal means of percentage returns from fitting the mixed effects model specified in the methods section, as well as the distribution of the returns for participants across all trials are presented in Figure \ref{fig:violinPanels}.A.-->

<!-- Investments averaged `r round(avg_pct$avg_inv,3)*100` $\%$ of endowment which is less than the $63.2\%$ documented by @fiedler_social_2011 and higher than the $37.3%$ documented by @charness_investment_2008, although the latter used repeated one shot games with random matching. Returns averaged `r round(avg_pct$avg_ret,3)*100` $\%$  of the total yield (3x investment sent), which is higher than the $35\%$ found in the same setting in @fiedler_social_2011.  -->

@@

```
```{r, include=FALSE, cache=TRUE}

@@

cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```
```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")

@@

<!-- If the intervention has an effect, then we should expect a higher increase in returns after the intervention compared to the control condition, which should be evident from an interaction between Condition (intervention vs control) and Game-number (pre- vs post-manipulation). Therefore, the variable of interest here is the interaction between Condition and Game-number.  -->

@@

Average investments and returns prior to the "defection round" (Figure \ref{fig:gamesPlot}) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. 
Mixed-effects analysis on the returns shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f`<!--$F(1,315.31) = 9.52$, $p = .002$-->, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f`<!--$F(1,314.2) = 26.9$, $p < .001$-->. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:violinPanels}.A). This indicates the intervention was effective in increasing cooperative behaviour. There was also a significant main effect of Investment, `r papaja::apa_print(mod_returns_pct)$full_result$inv_scaled`<!--$F(1,9208) = 373.6$, $p < 0.001$-->, such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, `r papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled`<!--$F(1,9207) = 45.38$, $p < .001$-->, indicates the positive effect of investment on percentage returns was greater in the control than intervention condition. <!-- This suggests that the higher returns we see in the intervention group are not simply due to higher investments.--> There was also an Investment by Game interaction, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled`. Finally, we find a three way interaction between Game, Condition and Investment, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled`<!--($F(1,8988)= 24.6, p < 0.001$),-->, showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game (pre- vs post intervention).

<!--
group had higher percentage returns than the control group.
We find an interaction effect between Condition (intervention vs. control) and Game-number (pre- vs. post-manipulation) ($F(1,314.2) = 26.9, p < 0.001$). Post-hoc tests show that there was an increase in the percentage returned in the intervention group (`r papaja::apa_print(prepost_bycond)$statistic$Intervention_Pre_post`) but not in the control group (`r papaja::apa_print(prepost_bycond)$statistic$Control_Pre_post`). We also found a significant main effect for Condition ( $F(1,315.31) = 9.52, p = 0.002)$), such that the intervention group had higher percentage returns than the control group. There was also a significant main effect of Investment( $F(1,9208) = 373.6, p < 0.001$) such that higher investments were associated with higher percentage returns.

There was an interaction effect of Investment by Condition ($F(1,9207) =45.38, p < 0.001$), such that the positive effect of investment on percentage returns was greater in the control group than in the intervention group. This suggests that the higher returns we see in the intervention group are not simply due to higher investments. Finally, we find a three way interaction between Game-number, Condition and Investment ($F(1,8988)= 24.6, p < 0.001$), showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game-number.

-->

@@

<!-- Taken together, we find that participants in the control condition sent lower percentage returns in the second game, despite the HMM investor sending on average higher investments in the second game. Those in the intervention group returned higher percentage returns in the second game, with the investor also sending higher investments. These higher returns in the intervention compared to the control condition were not purely driven by reciprocity towards higher investments, since we find a Condition by Game interaction whilst controlling for investment in the model, and a reduced effect of investment in the intervention condition. <!-- since we find this interaction effect whilst controlling for investments in our model of participants returns.-->

<!-- ## Looking at pre- and post-defection trials only -->

@@

We next analysed returns separately for rounds prior to the pre-programmed transgression (rounds 1 to 11 pre-intervention and 1 to 12 post-intervention) and rounds after (rounds 12 to 15 pre-intervention and rounds 13 to 15 post-intervention). Applying the same mixed effects model as before to returns before the transgression largely replicates the results over all trials. Participants in the intervention condition increased their returns in the second game, `r papaja::apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`. There was no evidence that participants in the control condition changed their returns in pre-transgression trials between the first and second RTG, `r papaja::apa_print(pairs_pre_int)$full_result$Control_Pre_post`. Full results are provided in the SI.

<!--
investments Analysis  and fit the same linear mixed effects model to percentage returns as the one for the full data. Estimated marginal means of percentage returns as well as the distribution of the returns for participants across all trials are presented in Figure \ref{fig:violinPanels}.B. We do find an interaction effect between Condition and Game number ($F(1,317.99) = 17.1, p < 0.001$). This was due to an increase in the percentage returned pre low-investment trial in the intervention condition (`r apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`) but not in the control group (`r apa_print(pairs_pre_int)$full_result$Control_Pre_post`). This indicates that participants were sending back higher returns in the intervention condition but not in the control condition even before the defection trial.
-->

@@

<!-- The intervention focused on repairing cooperation after a "transgressive action" by the investor. To assess whether the intervention succeeded, we next considered the returns by participants after the pre-programmed transgression (Figure \ref{fig:violinPanels}.D).--> <!-- We fit the same linear mixed effects model to percentage returns as the one for the full data while restricting the data to rounds 12 to 15 in the first game, and 13 to 15 in the second game. These trials represent all trials after the pre-programmed defection of the investor (low investment of $2$) until the end of the game (round 15).
-->
<!-- Figure \ref{fig:violinPanels}.D shows the marginal means of trustee returns and their distribution by Game-number and Condition.--> Applying the same model to the returns after the transgression (see SI for full results), we again find <!-- A mixed-model analysis shows similar results to those of the model applied to all returns (see SI for full results). There was again --> a significant interaction between Condition and Game. Participants in the control condition decreased their post-transgression returns from the first to the second RTG, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`. There was no significant change for participants in the intervention condition, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`. <!-- So, whilst the intervention did not appear to move participants to increase their returns after a transgression by the investor, it may have countered the decrease in returns showed by participants in the control condition. -->

@@

<!-- We found a main effect of Condition with returns on the intervention group higher than returns in the control group (`r apa_print(pairs(postcoax_bycond))$full_result`). We also found a main effect of Game-number, with returns higher for the pre-manipulation game compared to the post-manipulation game  (`r apa_print(pairs(postcoax_bygame))$full_result`). -->

@@

# colors <- c("Investment" = "blue", "Valence" = "red", "Arousal" = "green")
# 
# ggplot(datCoax,aes(x=roundNum)) + 
#   stat_summary(fun = "mean", geom = "line",group =1, aes(y=emo_scaled_x), color = "red") + 
#   stat_summary(fun = "mean", geom = "line",group =1, aes(y=emo_scaled_y), color = "green") + 
#   stat_summary(fun = "mean", geom = "line",group =1, aes(y=scale(investment)), color = "blue") + 
#   scale_color_manual(name = " ",values = colors, labels = c("Investment", "Emotion valence", "Emotion arousal")) +
#   labs(x = "Round", y="Series Z-score") +
#   geom_hline(yintercept = 0, linetype="dotted", color = "grey", linewidth=1.5) +
#   facet_grid(cols = vars(gameNum.f)) +
#   theme_bw() + 
#   theme(legend.position = "bottom")
#

# # BOTH IN ONE GRAPH 
# library(ggplot2)
# 
# # Define the colors in alphabetical order
# colors <- c("Arousal" = "green", "Investment" = "blue", "Valence" = "red")
# 
# # Plot 'pre'
# p <- ggplot(datCoax[datCoax$gameNum.f == 'pre', ], aes(x=roundNum)) + 
#   stat_summary(fun = "mean", geom = "line", group =1, aes(y=emo_scaled_x, color = "Valence", linetype = "pre")) + 
#   stat_summary(fun = "mean", geom = "line", group =1, aes(y=emo_scaled_y, color = "Arousal", linetype = "pre")) + 
#   stat_summary(fun = "mean", geom = "line", group =1, aes(y=inv_scaled, color = "Investment", linetype = "pre")) +
# # Plot 'post'
#   stat_summary(data = datCoax[datCoax$gameNum.f == 'post', ], fun = "mean", geom = "line", group =1, aes(x=roundNum, y=emo_scaled_x, color = "Valence", linetype = "post")) + 
#   stat_summary(data = datCoax[datCoax$gameNum.f == 'post', ], fun = "mean", geom = "line", group =1, aes(x=roundNum, y=emo_scaled_y, color = "Arousal", linetype = "post")) + 
#   stat_summary(data = datCoax[datCoax$gameNum.f == 'post', ], fun = "mean", geom = "line", group =1, aes(x=roundNum, y=inv_scaled, color = "Investment", linetype = "post")) +
#   labs(x = "Round", y="Series Z-score") +
#   geom_hline(yintercept = 0, linetype="dotted", color = "grey", linewidth=1.5) +
#   scale_color_manual(name = " ", values = colors, labels = c("Emotion arousal", "Investment", "Emotion valence")) +
#   scale_linetype_manual(values = c("pre" = "dotted", "post" = "solid"), labels = c("Pre", "Post")) +
#   theme_bw() + 
#   theme(legend.position = "bottom")
# 
# # Print the plot
# print(p)

@@

<!-- Figure \ref{fig:emoPlotCoax} shows the average, across participants *in the intervention condition*, of emotion valence, the degree of arousal as well as the HMM's investment z-scores. We find a significant positive correlation between the two emotion axis ($r=0.52, p < 0.001$) as well as between valence and Investment ($r=0.47, p < 0.001$), and arousal and Investment ($r=0.50, p < 0.001$). -->

@@

<!--
A key aspect of effective interventions is the emotional change. We examined emotion self-reports through fitting a linear mixed effects model to both emotion dimensions (valence and arousal) for the intervention group to explore whether the intervention changed the emotional appraisal of the investment, with fixed effects for Game-number (pre or post intervention) and Investment, as well as interaction between Investment and Game-number, with participant-wide random intercepts and random slopes for Game-number.

Higher investments were associated with more positive emotions ($F(1,4664) = 2902, p < 0.001$), and higher arousal ($F(1,4668.4) = 1919, p < 0.001$), while positive emotions declined between the two games ($\Delta M = 0.11, 95\% \ CI = [0.05, 0.18], t(\infty) = 3.72, p < .001$). However, there was no evidence that emotional reaction to the investment was lessened after the intervention. This means that participants in the intervention condition returned higher amounts post-intervention despite the emotional reaction to the lower investment being largely unchanged.
-->

<!-- For emotional valence, we find a main effect of Investment ($F(1,4664) = 2902, p < 0.001$), with higher investment leading to more positive emotions. We also find a main effect for Game-number, due to the reported emotional valence of investments decreasing significantly after the intervention ($\Delta M = 0.11, 95\% \ CI = [0.05, 0.18], t(\infty) = 3.72, p < .001$). For emotional arousal, fitting a similar mixed effects model we find only a main effect of Investment  ($F(1,4668.4) = 1919, p < 0.001$) with higher investment leading to higher arousal. These effects are still present if we exclude the defection trial. We did not find any interaction between emotional valence or arousal and the investment, which would indicate a moderating effect of the intervention on the reaction to investments. -->

@@

<!-- One possible reason of the differentiated returns post manipulation between conditions is that participants had a different perception of the investor, post-manipulation, in the control condition compared to the control condition. To explore this possibility, we can determine whether participants rated their artificial opponents differently between conditions (intervention vs control) and games (pre- vs post-intervention) using the players ratings they had to submit at the end of each game. To analyse the participants rating of their opponents, we used a linear mixed-effects model for each of the four players ratings (cooperativeness, selfishness, trustworthiness, friendliness) with fixed effects for Game-number (pre or post manipulation) and Condition (intervention or control), as well as interaction between condition and Game-number, and participant-wise random intercepts.  -->

@@

<!-- showed a main effect of Game-number on all attributes, with participants rating the second player as less cooperative (∆M = 0.42, 95% CI [0.15, 0.69], t(317.00) = 3.10, p = .002), less trustworthy (∆M = 0.43, 95% CI [0.16, 0.70], t(317.00) = 3.19, p = .002), less friendly (∆M = 0.40, 95% CI [0.17, 0.64], t(317.00) = 3.36, p = .001) and more selfish(∆M = −0.36, 95% CI [−0.61, −0.10], t(317.00) = −2.76, p = .006). Importantly, there was a main effect of Condition , with participants in the intervention condition rated players higher than those in the control condition on cooperativeness (∆M = 0.40, 95% CI [0.00, 0.80], t(317.00) = 1.95, p = .052) and lower on selfishness (∆M = −0.41, 95% CI [−0.80, −0.02], t(317.00) = −2.04, p = .042). There was no evidence for an interaction effect between Game-number and Condition on any of the attributes. -->

@@

<!--
Participants rated their artificial opponents differently between conditions (intervention vs control) and games (pre- vs post-intervention). The players ratings submitted at the end of each game showed a main effect of Game-number on all attributes, with participants rating the second player as less cooperative (`r apa_print(pairs(emmeans::emmeans(coop_mod, c("gameNum.f"))))$full_result`), less trustworthy (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("gameNum.f"))))$full_result`), less friendly (`r apa_print(pairs(emmeans::emmeans(friend_mod, c("gameNum.f"))))$full_result`) and more selfish(`r apa_print(pairs(emmeans::emmeans(self_mod, c("gameNum.f"))))$full_result`). We also find a main effect of Condition as participants in the intervention condition rated players higher than those in the control condition on cooperativeness (`r apa_print(pairs(emmeans::emmeans(trust_mod, c("condition.f"))))$full_result`) and lower on selfishness (`r apa_print(pairs(emmeans::emmeans(self_mod, c("condition.f"))))$full_result`). There was no evidence for an interaction effect between Game-number and Condition on any of the attributes.
-->

<!-- However, breaking down the interaction term by Condition, we find that participants rated the post-intervention player significantly lower on all positive attributes and higher on the negative one (selfish). There was no such difference in player ratings the control condition. -->

@@

We next asked whether the intervention had any discernible effect on participants' behaviour in a different, repeated Prisoner's Dilemma game. Predicting the probability of a cooperative action with a logistic mixed-effects regression model, with Condition and Phase (before or after defection trial) as fixed effects and a random intercept for participants, showed a decline in cooperation after defection by the other player, `r papaja::apa_print(ipd_mod)$full_result$phase`, but no evidence for a different cooperation rate in the intervention condition compared to the control condition, `r `r papaja::apa_print(ipd_mod)$full_result$condition_f`, or a different response to defection between the conditions, `r papaja::apa_print(ipd_mod)$full_result$condition_f_phase`. As such, there is no evidence that the intervention affected behaviour in this game. <!-- We also found no difference in cooperation rates post defection trial between conditions.-->

## Self report and debrief questionnaires

<!-- MS: This section doesn't say much. It could be reasonable to include questionnaire scores in the main analyses throughout (e.g. does BPD trait have an effect or moderate any other effects? IF not, a statement somewhere that the scales didn't have an effect, and showing results of accompanying analyses in SI would be fine -->

@@

```{r, include=FALSE}
# Install the 'psych' package if you haven't already
if (!requireNamespace("psych", quietly = TRUE)) {
  install.packages("psych")
}

@@

We find no interaction effect of the questionnaire scores with the Condition variable, nor a main effect of the questionnaire scores on participants returns in the RTG.

@@

When asked whether they thought their opponent was Human or not, $40$% of participants thought they were either facing a human or were not sure of the nature of the opponent. Many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial opponent's behaviour.

@@

# THE CODE BELOW WAS NOT USED; ALL CONVERGED. PUT IT HERE IN CASE IT IS NEEDED LATER ON.
# library(depmixS4)
# 
# load("data/fittedSimple.RData")
# load("data/fittedCtrl.RData")
# load("data/fittedCoax.RData")
# load("data/fittedFullCtrst.RData")
# load("data/fittedPrePost.RData")
# load("data/newFull.RData")
# 
# has_converged <- function(object, numeric = TRUE){
#   msg <- object@message
#   if(numeric) as.integer(grepl("converged", msg)) else 0
# }
# 
# 
# fit_until_convergence <- function(fitted_model, max_iterations=2) {
#   cat(fitted_model@nstates)
#   iterations <- 0
#   converged <- has_converged(fitted_model)
#   
#   while (!converged && iterations < max_iterations) {
#     # Use parameters from the fitted model as initial parameters
#     initial_params <- getpars(fitted_model)
#     new_model <- setpars(fitted_model, initial_params)
#     
#     # Refit the model
#     fitted_model <- fit(new_model)
#     converged <- has_converged(fitted_model)
#     iterations <- iterations + 1
#   }
#   
#   if (converged) {
#     cat("NS = ", fitted_model@nstates,"Convergence has been reached.\n")
#     flush.console()
#     return(fitted_model)
#   } else {
#     cat("Maximum iterations reached. Convergence has not been reached.\n")
#     flush.console()
#     return(fitted_model)
#   }
# }
# 
# 
# run_parallel_fitting <- function(mods_vec,num_cores = 5) {
#   # Set up the parallel backend
#   my.cluster <- parallel::makeCluster(num_cores, type = "PSOCK")
#   doParallel::registerDoParallel(cl = my.cluster)
#   
#   # Export the fit_until_convergence function to the workers
#   parallel::clusterExport(my.cluster, "fit_until_convergence")
#   parallel::clusterExport(my.cluster, "has_converged")
# 
#   # Run the loop in parallel using foreach
#   converged_mods_vec <- foreach(mod = mods_vec, .packages = c("depmixS4"), .combine = "c") %dopar% {
#     cat(mod@nstates)
#     fit_until_convergence(mod)
#   }
# 
#   # Stop the parallel backend
#   parallel::stopCluster(my.cluster)
#   return(converged_mods_vec)
# }
# 
# 
# sapply(fittedSimple[2:7],has_converged)
# sapply(fittedCtrl[2:7],has_converged)
# sapply(fittedCoax[2:7],has_converged)
# sapply(fittedFullCtrst[2:7],has_converged)
# sapply(fittedPrePost[2:7],has_converged)
# sapply(newFull[2:7],has_converged)

@@

@@

```{r postStatesBestMod,  echo=FALSE, fig.cap = "A: Distribution of participants' percentage return for each of the latent states in the 5 state HMM-full model. The latent states are ordered by the mean of the Gaussian that best fits the policy in that state, so higher numbered states are more pro-social. B: Distribution of posterior trustee states post manipulation by condition for all rounds, as estimated by the most likely posterior state in the best fitting HMM model (HMM-full) using a local decoding procedure.", fig.align="center", fig.height = 8}

>>>>

The strategy of the computerised investor was modelled on behaviour of human investors in a 10-round RTG <!-- reference to data -->.  Using this data, we estimated a hidden Markov model (HMM) on investors' behaviour with three latent states, reflecting "low-trust", "medium-trust", or "high-trust". Each latent state was associated with a distribution over all possible investor actions (from 0 to 20 investment) that reflected the amount of trust (Figure \ref{fig:HMMPanels}.A). Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of the net return in the previous round (see Figure \ref{fig:HMMPanels}.B).

In order to probe efforts to repair trust, the computerised agent was programmed to provide a low investment on round 12 pre-intervention and 13 post-intervention. On all other rounds, the agent's actions were derived from the HMM (disregarding the participant's response on pre-programmed low investment rounds).

@@

Participants played 7 rounds of a repeated Prisoner's Dilemma. In each round, participants could choose between: a cooperative action with a reward of 5 if the player also cooperated and a reward of 1 if not, or a non-cooperative action that would yield a 7 points if the other person chooses the cooperative action and 2 points if not. The Nash equilibrium for a single-round version is to choose the non-cooperative action.

The computerised agent was programmed to act according to a tit-for-tat strategy [@axelrod_evolution_1981], starting with a cooperative action and then mirroring what the other player chose in the preceding round. On round 4, the agent was pre-programmed to choose the defect action, regardless of the participant's preceding action.

@@

The intervention was built on interventions from DBT skills training, asking patients to reflect on the consequences of actions taken in emotional states [@linehan_dbt_2015]. Specifically, participants were presented with a low investment and asked to indicate their response. They were then invited to consider what their ultimate aim in the game was and whether this response was most likely to achieve their aim. The intervention is detailed in the supplementary information.

@@

At the start of the experiment (Figure \ref{fig:timeline}), participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Phase one was a 15 round repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor, participants were asked to provide feedback, with participants in the Intervention condition rating their feeling in terms of valence (from negative to positive) and arousal (from low to high), and participants in the Control condition rating the investment in terms of speed (from slow to fast) and magnitude (from low to high). These ratings were made by indicating a point on a two-dimensional grid. Participants then decided on how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, selfish, trustworthy and friendly they perceived their interaction partner (all on a scale from 1 to 10). After phase one, participants in the intervention condition completed the intervention, and participants in the control condition solved anagrams. Subsequent phase two was similar to phase one, with participants being told they would face a new player. Phase three consisted of 7 rounds of the repeated prisoner's dilemma game (RPD), with participants informed they would face a third player. Participants then completed questionnaires related to mentalising abilities, emotion regulation, and BPD traits (see the supplement for details). They were then asked about the strategy in the games, as well as whether they thought the other players were human or computer agents. They were then debriefed and thanked for their participation.

@@

@@

@@

@@

cond_effect <- pairs(emmeans::emmeans(mod_returns_pct, c("condition.f"), by="gameNum.f"))
```

```{r, include=FALSE}
pctRet_all <- afex::afex_plot(mod_returns_pct, x = "gameNum.f", trace = "condition.f", dodge = 0.8, error = "within",
            mapping = c("linetype", "shape", "fill"),
            data_geom = geom_violin, 
            data_arg = list(width = 0.5),
            factor_levels = list(gameNum.f = c("Pre", "Post")), legend_title = "Condition") + 
  theme_bw() + 
  labs(y = "Percentage Return", x = "Game") + 
  ggtitle("All rounds percentage returns") + 
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "bottom")

@@

@@

Average investments and returns prior to the "defection round" (Figure \ref{fig:gamesPlot}) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. 
Mixed-effects analysis on the returns shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f`<!--$F(1,315.31) = 9.52$, $p = .002$-->, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f`<!--$F(1,314.2) = 26.9$, $p < .001$-->. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post, `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:violinPanels}.A). This indicates the intervention was effective in increasing cooperative behaviour. There was also a significant main effect of Investment, `r papaja::apa_print(mod_returns_pct)$full_result$inv_scaled`<!--$F(1,9208) = 373.6$, $p < 0.001$-->, such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, `r papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled`<!--$F(1,9207) = 45.38$, $p < .001$-->, indicates the positive effect of investment on percentage returns was greater in the control than intervention condition. <!-- This suggests that the higher returns we see in the intervention group are not simply due to higher investments.--> There was also an Investment by Game interaction, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled`. Finally, we find a three way interaction between Game, Condition and Investment, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled`<!--($F(1,8988)= 24.6, p < 0.001$),-->, showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game (pre- vs post intervention).

@@

@@

We next analysed returns separately for rounds prior to the pre-programmed transgression (rounds 1 to 11 pre-intervention and 1 to 12 post-intervention) and rounds after (rounds 12 to 15 pre-intervention and rounds 13 to 15 post-intervention). Applying the same mixed effects model as before to returns before the transgression largely replicates the results over all trials. Participants in the intervention condition increased their returns in the second game, `r papaja::apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`. There was no evidence that participants in the control condition changed their returns in pre-transgression trials between the first and second RTG, `r papaja::apa_print(pairs_pre_int)$full_result$Control_Pre_post`. Full results are provided in the supplement.

@@

Applying the same model to the returns after the transgression (see the supplement for full results), we again find a significant interaction between Condition and Game. Participants in the control condition decreased their post-transgression returns from the first to the second RTG, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`. There was no significant change for participants in the intervention condition, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`. When questionnaire scores for borderline traits, mentalising and emotion regulation abilities were added as explanatory variables, we found no interaction effect of these questionnaire scores with the Condition variable, nor a main effect of the questionnaire scores on participants returns in the RTG.

@@

@@

@@

@@

@@

@@

@@

When asked during debrief whether they thought the investors they faced were Human or not, $40$% of participants thought they were either facing a human or were not sure of the nature of the opponent. Many answers reflected participants projecting human traits such as "spitefulness" or "greed" onto the artificial opponent's behaviour.

@@

We next asked whether the intervention had any discernible effect on participants' behaviour in a different, repeated Prisoner's Dilemma game. Predicting the probability of a cooperative action with a logistic mixed-effects regression model, with Condition and Phase (before or after defection trial) as fixed effects and a random intercept for participants, showed a decline in cooperation after defection by the other player, `r papaja::apa_print(ipd_mod)$full_result$phase`, but no evidence for a different cooperation rate in the intervention condition compared to the control condition, `r papaja::apa_print(ipd_mod)$full_result$condition_f`, or a different response to defection between the conditions, `r papaja::apa_print(ipd_mod)$full_result$condition_f_phase`. As such, there is no evidence that the intervention affected behaviour in this game. <!-- We also found no difference in cooperation rates post defection trial between conditions.-->

@@

```{r, include=FALSE}
# Install the 'psych' package
if (!requireNamespace("psych", quietly = TRUE)) {
  install.packages("psych")
}

@@

@@

@@

has_converged <- function(object, numeric = TRUE){
  msg <- object@message
  if(numeric) as.integer(grepl("converged", msg)) else 0
}

sapply(fittedSimple[2:7],has_converged)
sapply(fittedCtrl[2:7],has_converged)
sapply(fittedCoax[2:7],has_converged)
sapply(fittedFullCtrst[2:7],has_converged)
sapply(fittedPrePost[2:7],has_converged)
sapply(newFull[2:7],has_converged)

@@

```{r}

# THE CODE BELOW WAS NOT USED; ALL CONVERGED. PUT IT HERE IN CASE IT IS NEEDED LATER ON.
# library(depmixS4)
# fit_until_convergence <- function(fitted_model, max_iterations=2) {
#   cat(fitted_model@nstates)
#   iterations <- 0
#   converged <- has_converged(fitted_model)
#   
#   while (!converged && iterations < max_iterations) {
#     # Use parameters from the fitted model as initial parameters
#     initial_params <- getpars(fitted_model)
#     new_model <- setpars(fitted_model, initial_params)
#     
#     # Refit the model
#     fitted_model <- fit(new_model)
#     converged <- has_converged(fitted_model)
#     iterations <- iterations + 1
#   }
#   
#   if (converged) {
#     cat("NS = ", fitted_model@nstates,"Convergence has been reached.\n")
#     flush.console()
#     return(fitted_model)
#   } else {
#     cat("Maximum iterations reached. Convergence has not been reached.\n")
#     flush.console()
#     return(fitted_model)
#   }
# }
# 
# 
# run_parallel_fitting <- function(mods_vec,num_cores = 5) {
#   # Set up the parallel backend
#   my.cluster <- parallel::makeCluster(num_cores, type = "PSOCK")
#   doParallel::registerDoParallel(cl = my.cluster)
#   
#   # Export the fit_until_convergence function to the workers
#   parallel::clusterExport(my.cluster, "fit_until_convergence")
#   parallel::clusterExport(my.cluster, "has_converged")
# 
#   # Run the loop in parallel using foreach
#   converged_mods_vec <- foreach(mod = mods_vec, .packages = c("depmixS4"), .combine = "c") %dopar% {
#     cat(mod@nstates)
#     fit_until_convergence(mod)
#   }
# 
#   # Stop the parallel backend
#   parallel::stopCluster(my.cluster)
#   return(converged_mods_vec)
# }
# 
```

@@

```{r postStatesBestModOld,  echo=FALSE, fig.cap = "A: Distribution of participants' percentage return for each of the latent states in the 5 state HMM-full model. The latent states are ordered by the mean of the Gaussian that best fits the policy in that state, so higher numbered states are more pro-social. B: Distribution of posterior trustee states post manipulation by condition for all rounds, as estimated by the most likely posterior state in the best fitting HMM model (HMM-full) using a local decoding procedure.", fig.align="center", fig.height = 8}

# ggpubr::ggarrange(trusteeHMMPlot , postHMM , nrow=2, ncol=1,labels = c("A", "B"),common.legend = FALSE)
